Resource Buying Games

### 2017-03-02 slides.tex
a few small corrections
started work on paper
last slide polish

\documentclass{beamer}
\usepackage{pgfplots}
\usepackage{tikz}
\usepackage{csquotes}
\usepackage{cancel}
\usepackage{stmaryrd}
\usepackage{url}
\usetheme{metropolis}
\title{Resource Buying Games}
\author[C. Welzel]{Christoph Welzel}

@@

\begin{frame}
  \frametitle{Introduction}
    \begin{itemize}
      \item Resources: Nodes/edges in a graph, Machines to schedule jobs on,
        \dots
      \item Players: jointly buy resources to meet individual
        (potentially overlapping) demands
      \item<2-> Every player contributes to buy certain resources
      \item<3-> Payoff for a player depends on what she payed to meet her demands
        of resources
    \end{itemize}
    \uncover<4->{\begin{definition}[Pure Nash Equilibrium(informal)]
      A \emph{Pure Nash Equilibrium (PNE)} is given if for every player holds
      that, given all other players stick to their strategies, she cannot alter
      her behaviour to achieve a better payoff for herself.
    \end{definition}}
\end{frame}

\section{Basic Definitions}
\begin{frame}
  \frametitle{Definition: congestion model}
  \begin{equation*}
    \mathcal{M} = \tupel{N, E, \mathcal{S}, \left(d_{i}\right)_{i\in N},
    \left(c_{e}\right)_{e\in E}}
  \end{equation*}
  \vspace{-1cm}
  \begin{itemize}
    \item $N$: set of players
    \item<2-> $E$: set of resources
    \item<3-> $\mathcal{S} = \times_{i\in N}\mathcal{S}_{i}$ with $\mathcal{S}_{i}$
      set of desired resources $S_{i}\subseteq E$: set of configurations
    \item<4-> $d_{i}$: demand of player $i$ for her resource
      \begin{itemize}
        \item<4-> $d_{i} = 1$ for an unweighted game
        \item<4-> $\ell_{e}(S) = \sum_{i\in N:e\in S_{i}}d_{i}$: load of $e$
          in configuration $S\in\mathcal{S}$
      \end{itemize}
    \item<5-> $c_{e}:\mathbb{N}\rightarrow\mathbb{N}$ with
      $c_{e}(S) = c_{e}(\ell_{e}(S))$: cost for the resource $e$ under the load
      of $S$
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Definition: weighted resource buying game}
  \begin{equation*}
    \mathcal{M} = \tupel{N, E, \mathcal{S}, \left(d_{i}\right)_{i\in N},
    \left(c_{e}\right)_{e\in E}}
  \end{equation*}
  \begin{equation*}
    \leadsto\mathcal{G} = \tupel{N, \mathcal{S}\times\mathcal{P}, \pi}
  \end{equation*}
  \vspace{-1cm}
  \begin{itemize}
    \item<2-> $\mathcal{P}=\times_{i\in N} P_{i}$ with
      $P_{i} = \mathbb{R}^{|E|}_{+}$: payment of player $i$
    \item<3-> $e\in E$ is bought under $S\in \mathcal{S}$ if
      $\sum_{i\in N}p_{i}^{e} \geq c_{e}(S)$
    \item<4-> $\pi_{i}(S) =
      \begin{cases}
        \sum_{e\in E}p_{i}^{e} &\textit{if }S_{i}\textit{ is bought}\\
        \infty &\textit{otherwise}
      \end{cases}$: {private cost of player $i$}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{A few simplifications}
  \begin{itemize}
    \item \emph{unweighted} games
    \item<2-> marginally non-increasing cost-function $c_{e}$:
      \begin{equation*}
        c_{e}(x + \delta) - c_{e}(x) \geq c_{e}(y + \delta) - c_{e}(y)
        \;\;x\leq y, \delta \in \mathbb{N}
      \end{equation*}
      \begin{center}
        \begin{tikzpicture}
          \begin{axis}[
              xlabel=$x$,
              ylabel=$c_{e}(x)$,
              smooth,
              width=5cm,
              height=3cm,
              ticks=none,
              no markers
            ]
              \addplot{ln(x)};
          \end{axis}
        \end{tikzpicture}
      \end{center}
    \item<3-> restrict players demands to a specific mathematical structure:
      Matroids
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Operations on Matroids $\mathcal{M} = \tupel{E, \mathcal{B}}$}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{center}
      Contraction:
    \end{center}
    \vspace{-0.6cm}
    \begin{align*}
      &\mathcal{M}/e = \tupel{E\setminus\set{e}, \mathcal{B}/e}\text{ with}\\
      &\mathcal{B}/e = \set{B\subseteq (E\setminus\set{e})\middle| B\cup\set{e}\in\mathcal{B}}
    \end{align*}
    \column<4->{0.5\textwidth}
    \begin{center}
      Deletion:
    \end{center}
    \vspace{-0.6cm}
    \begin{align*}
      &\mathcal{M}\setminus e = \tupel{E\setminus\set{e},\mathcal{B}\setminus \set{e}}\text{ with}\\
      &\mathcal{B}\setminus e = \set{B\subseteq E\setminus\set{e}\middle|B\in\mathcal{B}}
    \end{align*}
  \end{columns}
  \begin{center}
    \begin{tikzpicture}
      \uncover<2->{\node (G) {\input{tikz/graphp1}};
      \node[right=1 of G] (leadsto) {\Huge$\leadsto$};}
      \uncover<3>{\node[right=1 of leadsto] (G') {\input{tikz/graphp1Cont}};}
      \uncover<5>{\node[right=1 of leadsto] (G'') {\input{tikz/graphp1Del}};}
    \end{tikzpicture}
  \end{center}
\end{frame}

@@

\begin{frame}
  \frametitle{Soundness}
  \begin{itemize}
    \item Termination:
      \begin{itemize}
        \item<2-> inductively applying contraction operations lead to basis
        \item<3-> deletion cannot \enquote{loose} bases for other players since
          chosen cut is inclusion-wise minimal
      \end{itemize}
    \item Correctness
      \begin{itemize}
        \item<4-> Lemma: cost of the chosen element decreases monotonically in each
          iteration
        \item<5-> assume there is a beneficial alternative basis
          $\hat{B} = (B_{i}\setminus\set{g})\cup{f}$
          \begin{itemize}
            \item<6-> consider iteration $k$ in which $g$ is added to basis of $i$:
            \item<7-> if $f$ is in the corresonponding cut its cost never changes
              since it is deleted from every matroid $\rightarrow$ $f$ is not
              in the cut
            \item<8-> but $f$ and $g$ \enquote{open} same dimensionality, thus $f$
              is not left in contracted matroid of $i$
            \item<9-> $f$ was deleted before (in this iteration the cost of the
              chosen element was lower than for $f$ in contradiction to lemma)
          \end{itemize}
      \end{itemize}
  \end{itemize}
\end{frame}

\section{Weighted Matroid Games}
\begin{frame}
  \frametitle{Recall: Congestion Model}
  \begin{equation*}
    \mathcal{M} = \tupel{N, E, \mathcal{B}, \left(d_{i}\right)_{i\in N},
    \left(c_{e}\right)_{e\in E}}
  \end{equation*}
  \vspace{-1cm}
  \begin{itemize}
    \item before: $d_{i} = 1$ for all $i$
    \item<2-> now: $d_{i}\in\mathbb{N}_{>0}$ 
    \item<3-> auxiliary definitions:
      \begin{itemize}
        \item<4-> $N_{e}\tupel{B} = \set{i\in N\middle|e\in B_{i}}$
        \item<5-> for a fixed $B$: $\ex_{i}\tupel{e} =
          \set{f\in E\setminus\set{e}\middle|
          (B_{i}\setminus{e})\cup\set{f}\in\mathcal{B}_{i}}$: set of resources
          that can replace $e$ for player $i$
        \item<6-> $F = \bigcup_{i\in N}\bigcap_{B\in \mathcal{B}_{i}} B$:
          set of obligatory elements for at least one player
        \item<7-> $\Delta_{i}\tupel{B;e\rightarrow f} =
          c_{f}\tupel{\ell_{f}\tupel{B_{i}\cup\set{f}\setminus\set{e}}} -
          c_{f}\tupel{\ell_{f}\tupel{B}}$: marginal cost to buy resource $f$ if
          $i$ switches
        \item<8-> $\Delta_{i}^{e} = \min_{f\in\ex_{i}(e)}\set{\Delta_{i}(B;e\rightarrow f)}$:
          minimal marginal cost for exchanging $e$
      \end{itemize}
  \end{itemize}
\end{frame}

>>>>

\documentclass{beamer}
\usepackage{pgfplots}
\usepackage{tikz}
\usepackage{csquotes}
\usepackage{cancel}
\usepackage{stmaryrd}
\usepackage{url}
\usetheme{metropolis}
\title{Resource Buying Games}
\author[C. Welzel]{Christoph Welzel}
\date{March 3, 2017}

@@

\begin{frame}
  \frametitle{Introduction}
    \begin{itemize}
      \item Resources: Nodes/edges in a graph, Machines to schedule jobs on,
        \dots
      \item Players: jointly buy resources to meet individual
        (potentially overlapping) demands
      \item<2-> Every player contributes to buy certain resources
      \item<3-> Payoff for a player depends on what she payed to meet her demands
    \end{itemize}
    \uncover<4->{\begin{definition}[Pure Nash Equilibrium(informal)]
      A \emph{Pure Nash Equilibrium (PNE)} is given if for every player holds
      that, given all other players stick to their strategies, she cannot alter
      her behaviour to achieve a better payoff for herself.
    \end{definition}}
\end{frame}

\section{Basic Definitions}
\begin{frame}
  \frametitle{Definition: Congestion Model}
  \begin{equation*}
    \mathcal{M} = \tupel{N, E, \mathcal{S}, \left(d_{i}\right)_{i\in N},
    \left(c_{e}\right)_{e\in E}}
  \end{equation*}
  \vspace{-1cm}
  \begin{itemize}
    \item $N$: set of players
    \item<2-> $E$: set of resources
    \item<3-> $\mathcal{S} = \times_{i\in N}\mathcal{S}_{i}$ with $\mathcal{S}_{i}$
      set of desired resources $S_{i}\subseteq E$: set of configurations
    \item<4-> $d_{i}$: demand of player $i$ for her resource
      \begin{itemize}
        \item<4-> $d_{i} = 1$ for an unweighted game
        \item<4-> $\ell_{e}(S) = \sum_{i\in N:e\in S_{i}}d_{i}$: load of $e$
          in configuration $S\in\mathcal{S}$
      \end{itemize}
    \item<5-> $c_{e}:\mathbb{N}\rightarrow\mathbb{R}_{+}$ with
      $c_{e}(S) = c_{e}(\ell_{e}(S))$: cost for the resource $e$ under the load
      of $S$
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Definition: Weighted Resource Buying Game}
  \begin{equation*}
    \mathcal{M} = \tupel{N, E, \mathcal{S}, \left(d_{i}\right)_{i\in N},
    \left(c_{e}\right)_{e\in E}}
  \end{equation*}
  \begin{equation*}
    \leadsto\mathcal{G} = \tupel{N, \mathcal{S}\times\mathcal{P}, \pi}
  \end{equation*}
  \vspace{-1cm}
  \begin{itemize}
    \item<2-> $\mathcal{P}=\times_{i\in N} P_{i}$ with
      $P_{i} = \mathbb{R}^{|E|}_{+}$: payment of player $i$
    \item<3-> $e\in E$ \emph{is bought under $S\in \mathcal{S}$} if
      $\sum_{i\in N}p_{i}^{e} \geq c_{e}(S)$
    \item<4-> $\pi_{i}(S) =
      \begin{cases}
        \sum_{e\in E}p_{i}^{e} &\textit{if }S_{i}\textit{ is bought}\\
        \infty &\textit{otherwise}
      \end{cases}$: {private cost of player $i$}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Restricted Setup}
  \begin{itemize}
    \item \emph{unweighted} games
    \item<2-> marginally non-increasing cost-function $c_{e}$:
      \begin{equation*}
        c_{e}(x + \delta) - c_{e}(x) \geq c_{e}(y + \delta) - c_{e}(y)
        \;\;x\leq y, \delta \in \mathbb{N}
      \end{equation*}
      \begin{center}
        \begin{tikzpicture}
          \begin{axis}[
              xlabel=$x$,
              ylabel=$c_{e}(x)$,
              smooth,
              width=5cm,
              height=3cm,
              ticks=none,
              no markers
            ]
              \addplot{ln(x)};
          \end{axis}
        \end{tikzpicture}
      \end{center}
    \item<3-> restrict players demands to a specific mathematical structure:
      Matroids
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Operations on Matroid $\mathcal{M} = \tupel{E, \mathcal{B}}$}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{center}
      Contraction:
    \end{center}
    \vspace{-0.6cm}
    \begin{align*}
      &\mathcal{M}/e = \tupel{E\setminus\set{e}, \mathcal{B}/e}\text{ with}\\
      &\mathcal{B}/e = \set{B\subseteq (E\setminus\set{e})\middle| B\cup\set{e}\in\mathcal{B}}
    \end{align*}
    \column<4->{0.5\textwidth}
    \begin{center}
      Deletion:
    \end{center}
    \vspace{-0.6cm}
    \begin{align*}
      &\mathcal{M}\setminus e = \tupel{E\setminus\set{e},\mathcal{B}\setminus \set{e}}\text{ with}\\
      &\mathcal{B}\setminus e = \set{B\subseteq E\setminus\set{e}\middle|B\in\mathcal{B}}
    \end{align*}
  \end{columns}
  \begin{center}
    \begin{tikzpicture}
      \uncover<2->{\node (G) {\input{tikz/graphp1}};
      \node[right=1 of G] (leadsto) {\Huge$\leadsto$};}
      \uncover<3>{\node[right=1 of leadsto] (G') {\input{tikz/graphp1Cont}};}
      \uncover<5>{\node[right=1 of leadsto] (G'') {\input{tikz/graphp1Del}};}
    \end{tikzpicture}
  \end{center}
\end{frame}

@@

\begin{frame}
  \frametitle{Soundness}
  \begin{itemize}
    \item Termination:
      \begin{itemize}
        \item<2-> inductively applying contraction operations lead to basis
        \item<3-> deletion cannot \enquote{loose} bases for other players since
          chosen cut is inclusion-wise minimal
      \end{itemize}
    \item Correctness
      \begin{itemize}
        \item<4-> Lemma: cost of the chosen element decreases monotonically in each
          iteration
        \item<5-> assume there is a beneficial alternative basis
          $\hat{B} = (B_{i}\setminus\set{g})\cup{\set{f}}$
          \begin{itemize}
            \item<6-> consider iteration $k$ in which $g$ is added to basis of $i$:
            \item<7-> if $f$ is in the corresonponding cut its cost never changes
              since it is deleted from every matroid $\rightarrow$ $f$ is not
              in the cut
            \item<8-> but $f$ and $g$ \enquote{open} same dimensionality, thus $f$
              is not left in contracted matroid of $i$
            \item<9-> $f$ was deleted before (in this iteration the cost of the
              chosen element was lower than for $f$ in contradiction to lemma)
          \end{itemize}
      \end{itemize}
  \end{itemize}
\end{frame}

\section{Weighted Matroid Games}
\begin{frame}
  \frametitle{Recall: Congestion Model}
  \begin{equation*}
    \mathcal{M} = \tupel{N, E, \mathcal{B}, \left(d_{i}\right)_{i\in N},
    \left(c_{e}\right)_{e\in E}}
  \end{equation*}
  \vspace{-1cm}
  \begin{itemize}
    \item before: $d_{i} = 1$ for all $i$
    \item<2-> now: $d_{i}\in\mathbb{N}_{>0}$ 
    \item<3-> auxiliary definitions:
      \begin{itemize}
        \item<4-> $N_{e}\tupel{B} = \set{i\in N\middle|e\in B_{i}}$
        \item<5-> for a fixed $B$: $\ex_{i}\tupel{e} =
          \set{f\in E\setminus\set{e}\middle|
          (B_{i}\setminus\set{e})\cup\set{f}\in\mathcal{B}_{i}}$: set of resources
          that can replace $e$ for player $i$
        \item<6-> $F = \bigcup_{i\in N}\bigcap_{B\in \mathcal{B}_{i}} B$:
          set of obligatory elements for at least one player
        \item<7-> $\Delta_{i}\tupel{B;e\rightarrow f} =
          c_{f}\tupel{\ell_{f}\tupel{B_{i}\cup\set{f}\setminus\set{e}}} -
          c_{f}\tupel{\ell_{f}\tupel{B}}$: marginal cost to buy resource $f$ if
          $i$ switches
        \item<8-> $\Delta_{i}^{e} = \min_{f\in\ex_{i}(e)}\set{\Delta_{i}(B;e\rightarrow f)}$:
          minimal marginal cost for exchanging $e$
      \end{itemize}
  \end{itemize}
\end{frame}
